Annual Analysis of Delays in TTC Streetcars for the year 2023

### 2024-09-27 paper/paper.qmd
Update ReadMe, scripts for test and simulation and llm usage. Recompile paper and qmd document

## Data preprocessing and cleaning
The raw data is undergone a series of data cleaning steps to ensure adequate usability for the study. We first use the Janitor (@janitor) package to clean the data columns removing any spaces between them and using a consistent naming stype for them. Next, we proceed to split the date field into year and month fields and split the time field extracting the hour of day from it. Following this, we create a new column to denote the season based on the month of the year from the data. This helps during the analysis stage when we need to make inferences about the seasonal variation in average delays in streetcars.

@@

As a result of the above described data cleaning process, we select our desired features from the dataset. The following table shows the first five rows of our data and the selected features.

```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false

table_features <- gt(limited_row_data)
table_features |> 
  table_styling() |>
  tab_header(
    title = "Selected Features from TTC Streetcar Delay Data"
    ) |>
  tab_style(
    style = cell_text(
      weight = 'bold',
      font = google_font('Merriweather')
    ),
    locations = cells_title(group = 'title')
  ) |>
  tab_style(
    style = cell_fill(color = 'grey90'),
    locations = cells_body(rows = seq(1, nrow(limited_row_data), 2))
  )
cat("Table 1: First first five rows in the cleaned dataset with selected features")

## Measurement of units for selected features

@@

The following table describes the data types for the selected features in our study from the cleaned dataset after the necessary transformations.
```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false

# Combine into a table (data frame)
table <- data.frame(Feature = features, DataType = datatypes)
table_feature_datatypes <- gt(table)
table_feature_datatypes |>
  tab_header(
    title = "Selected Features and their Data Types"
    ) |>
  tab_style(
    style = cell_text(
      weight = 'bold',
      font = google_font('Merriweather')
    ),
    locations = cells_title(group = 'title')
  ) |>
  tab_style(
    style = cell_text(weight = 'bold'),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = 'grey90'),
    locations = cells_body(rows = seq(1, nrow(table), 2))
  )
  cat("Table 2: Datatypes for the selected features from the cleaned dataset.")
```

>>>>

\newpage
## Data preprocessing and cleaning

```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false
#| label: tbl-dataset-preview
#| tbl-cap: "First first five rows in the cleaned dataset with selected features"

table_features <- gt(limited_row_data)
table_features |> 
  table_styling() |>
  tab_header(
    title = "Selected Features from TTC Streetcar Delay Data"
    ) |>
  tab_style(
    style = cell_text(
      weight = 'bold',
      font = google_font('Merriweather')
    ),
    locations = cells_title(group = 'title')
  ) |>
  tab_style(
    style = cell_fill(color = 'grey90'),
    locations = cells_body(rows = seq(1, nrow(limited_row_data), 2))
  )

The raw data is undergone a series of data cleaning steps to ensure adequate usability for the study. We first use the Janitor (@janitor) package to clean the data columns removing any spaces between them and using a consistent naming stype for them. Next, we proceed to split the date field into year and month fields and split the time field extracting the hour of day from it. Following this, we create a new column to denote the season based on the month of the year from the data. This helps during the analysis stage when we need to make inferences about the seasonal variation in average delays in streetcars.

@@

As a result of the above described data cleaning process, we select our desired features from the dataset. @tbl-dataset-preview shows the first five rows of our data and the selected features.

\newpage
## Measurement of units for selected features

```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false
#| label: tbl-selected-features
#| tbl-cap: "Datatypes for the selected features from the cleaned dataset"

# Combine into a table (data frame)
table <- data.frame(Feature = features, DataType = datatypes)
table_feature_datatypes <- gt(table)
table_feature_datatypes |>
  tab_header(
    title = "Selected Features and their Data Types"
    ) |>
  tab_style(
    style = cell_text(
      weight = 'bold',
      font = google_font('Merriweather')
    ),
    locations = cells_title(group = 'title')
  ) |>
  tab_style(
    style = cell_text(weight = 'bold'),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = 'grey90'),
    locations = cells_body(rows = seq(1, nrow(table), 2))
  )
```

@@

@tbl-selected-features describes the data types for the selected features in our study from the cleaned dataset after the necessary transformations.
